### 2025-03-03 paper.tex
keep on writing

\documentclass[fontsize=11pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{color}
\usepackage{graphicx}
\usepackage{float}
\usepackage{polynom}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{gensymb}
\usepackage{tikz}
\tikzset{every picture/.style={font=\normalsize}}
\DeclareOldFontCommand{\rm}{\normalfont\rmfamily}{\mathrm}
\DeclareOldFontCommand{\it}{\normalfont\rmfamily}{\mathrm}
\usepackage{mathtools}
\usepackage{hyperref}
\hypersetup{colorlinks=false,
            pdfborder={0 0 0}}
\renewcommand{\thefootnote}{\alph{footnote}}
\newcommand{\R}{\mathbb{R}}
\usepackage{amsmath, bm}   % Extra math commands and environments from the AMS
\usepackage{amssymb}   % Special symbols from the AMS
\usepackage{amsthm}    % Enhanced theorem and proof environments from the AMS
\usepackage{latexsym}  % A few extra LaTeX symbols
\usepackage{rxn}
\usepackage[version=4]{mhchem}
\usepackage{cleveref}
\usepackage{url}
\providecommand{\email}{}
\renewcommand{\email}[1]{\texttt{#1}}
\usepackage{caption}
\captionsetup{labelfont=bf, labelsep=period, labelformat=simple, labelsep=colon, figurename=Fig.}
\usepackage[authoryear, round]{natbib}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{property}[theorem]{Property}

%\newcommand{\tL}{\tilde{L}}

\graphicspath{{./}{./figures/}{./figures/paper/}}

\setcounter{topnumber}{2}              %% 2
\setcounter{bottomnumber}{1}           %% 1
\setcounter{totalnumber}{3}            %% 3
\renewcommand{\topfraction}{0.9}       %% 0.7
\renewcommand{\bottomfraction}{0.9}    %% 0.3
\renewcommand{\textfraction}{0.1}      %% 0.2
\renewcommand{\floatpagefraction}{.7}  %% 0.5

\title{\bf{Neural Ordinary Differential Equations}}
\author{Prithvi Dake\\
Department of Chemical Engineering, University of California, Santa Barbara}
\date{\textit{ME255NN: Neural Networks for modeling, optimization and control} \\
\textit{Winter 2025}}

\begin{document}

\maketitle
\section{Introduction}

Conventional neural network architectures like RNNs and residual network
resemble explicit Euler scheme to solve ODEs. Thus NODEs can be viewed as 
the \textit{continuous} versions of such \textit{discretized} architectures.
Now one problem with the continuous version is propagating the gradients during
back-propagation while training. One way to solve this problem is augmenting
the ODE with its sensitivity equations. Such a scheme saves precious memory that
goes in storing intermediate gradients in brute-force back-prop. Thus, after
overcoming the training problem, we can extend the NODE architecture to other 
applications like the continuous version of normalizing flow and state-space models.

Literature review (6 articles atleast)

The report is structured as follows: We explain conventional NN architectures like
RNN and ResNet and connect it to its continuous version called as NODE. We also show
how we can handle back-prop for NODE. We also discuss continuous normalizing flows and
NODE-based state-space models followed by some interesting case-studies. We finally 
end the report with some critical conclusions and review of current developments.

\section{NODEs and its application}
\section{Case studies}
\section{Conclusions}

\newpage
\FloatBarrier
\bibliographystyle{abbrvnat}
\bibliography{abbreviations,articles,books,unpub,proceedings}
\end{document}